Policy optimization for dynamic spatiotemporal systems \hspace{1em} \insertframenumber

### 2014-08-02 pot/LaTeX/jsm2014.tex
presentation updates

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Assume observed data $Y_t(\bs)$ come from a skew-$t$ (Zhang and El-Shaarawi, 2012)
    \begin{align*}
      Y_t(\bs) = X_t(\bs)\beta + \alpha z_t + v_t(\bs)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $\alpha \in \calR$ controls the skew
      \item $z_t \iid N_{(0, \infty)}(0, \sigma^2_t)$ is a random effect
      \item $v_t(\bs)$ is a Gaussian process with variance $\sigma^2_t$ and \Matern correlation
      \item $\sigma_t^2 \iid \text{IG}(a, b)$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Skew-normal distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Conditioned on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
    \begin{itemize}
      \item After marginalizing over $\sigma^2_t$, this becomes a multivariate $t$-distribution with $2a$ degrees of freedom.
    \end{itemize}
    \item Can use standard geostatistical methods to fit this model.
    \item Predictions can be made through kriging.
    \item Marginalizing over $z_t$ and $\sigma^2_t$ (via MCMC),
    \begin{align*}
      Y_t(\bs) \sim \text{skew-t}(\mu, \alpha, \text{df}=2a)
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Random daily partition}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site.
    \begin{align*}
      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + v_t(\bs)
    \end{align*}
    \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
    $\{P_{t1}, \ldots, P_{tK}\}$ such that
    \begin{align*}
      P_{tk} = \{s : k = \argmin_\ell|| \bs - w_{t\ell}|| \}
    \end{align*}
    \item For $\bs \in P_{tk}$
    \begin{align*}
      z_t(\bs) &= z_{tk}\\
      \sigma^2_t(\bs) &= \sigma^2_{tk}
    \end{align*}
    \item Within each partition $Y_t(\bs)$ has the same distribution as before.
  \end{itemize}
\end{frame}

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Three main steps:
    \begin{enumerate}[1.]
      \item Impute missing observations and data below $T$
      \item Update parameters with random walk Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 9 different analysis methods incorporating
    \begin{itemize}
      \item Gaussian vs $t$
      \item 1 partition vs 5 partitions
      \item No thresholding vs thresholded
      \begin{itemize}
        \item Thresholded data at 0.90 sample quantile
      \end{itemize}
    \end{itemize}
    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
    \item Compare quantile and Brier scores using 5-fold cross validation
    \item Mean function modeled using a first-order spatial trend
  \end{itemize}
\end{frame}

\begin{frame}{Cross-validation results}
  \begin{table}[htbp]
    \small
    \centering
    \begin{tabular}{lrrrrrr}
           & \multicolumn{6}{c}{Quantile}\\
           \cline{2-7}
    & 0.900 & 0.950 & 0.980 & 0.990 & 0.995 & 0.999\\
  \hline
Gaussian & 16.38 & 15.76 & 15.02 & 14.52 & 14.08 & 13.22\\
$t$-1 ($T=0$) & 16.15 & 15.51 & 14.62 & 14.00 & 13.43 & 12.32\\
$t$-5 ($T=0$) & 13.61 & 12.66 & 11.61 & 10.96 & 10.40 & 9.34\\
$t$-1 ($T=0.9$) & 5.52  & 3.58  & 2.28  & 1.77  & 1.47  & 1.10\\
$t$-5 ($T=0.9$) & 5.98  & 4.27  & 2.99  & 2.41  & 2.03  & 1.49\\
skew $t$-1 ($T=0$) & 9.24  & 7.27  & 5.26  & 4.13  & 3.27  & 1.96\\
skew $t$-1 ($T=0.9$) & {\bf 4.91}  & {\bf 3.16} & {\bf 1.94}  & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
skew $t$-5 ($T=0$) & 15.81 & 14.46 & 12.74 & 11.57 & 10.57 & 8.60\\
skew $t$-5 ($T=0.9$) & & & & & &\\
\hline
    \end{tabular}
  \end{table}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Quantile scores -- Still waiting on last set
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Assume observed data $Y_t(\bs)$ come from a skew-$t$ (Zhang and El-Shaarawi, 2012)
    \begin{align*}
      Y_t(\bs) = X_t(\bs)\beta + \alpha z_t + v_t(\bs)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $\alpha \in \calR$ controls the skewness
      \item $z_t \iid N_{(0, \infty)}(0, \sigma^2_t)$ is a random effect
      \item $v_t(\bs)$ is a Gaussian process with variance $\sigma^2_t$ and \Matern correlation
      \item $\sigma_t^2 \iid \text{IG}(a, b)$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
    \item Can use standard geostatistical methods to fit this model.
    \item Predictions can be made through kriging.
    \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
    \begin{align*}
      Y_t(\bs) \sim \text{skew-t}(\mu, \alpha, \text{df}=2a)
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Random daily partition}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site.
    \begin{align*}
      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + \sigma(\bs) v_t(\bs)
    \end{align*}
    \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
    $P_{t1}, \ldots, P_{tK}$ such that
    \begin{align*}
      P_{tk} = \{s : k = \argmin_\ell|| \bs - w_{t\ell}|| \}
    \end{align*}
    \item For $\bs \in P_{tk}$
    \begin{align*}
      z_t(\bs) &= z_{tk}\\
      \sigma^2_t(\bs) &= \sigma^2_{tk}
    \end{align*}
    \item Within each partition $Y_t(\bs)$ has the same MVT distribution as before.
  \end{itemize}
\end{frame}

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Three main steps:
    \begin{enumerate}[1.]
      \item Impute missing observations and data below $T$
      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 9 different analysis methods incorporating
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
      \item $K=1$ partition vs $K=5$ partitions
      \item No thresholding vs thresholded
      \begin{itemize}
        \item Thresholded data at $T=0.90$ sample quantile
      \end{itemize}
    \end{itemize}
    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
    \item Compare quantile and Brier scores using 5-fold cross validation
    \item Mean function modeled using a first-order spatial trend
  \end{itemize}
\end{frame}

\begin{frame}{Cross-validation results}
  \begin{table}[htbp]
    \small
    \centering
    \begin{tabular}{l|c|r|rrrrr}
          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{Quantile}\\
           \hline
  Marginal & $K$ & $T$  & 0.900 & 0.950 & 0.990 & 0.995 & 0.999\\
  \hline
Gaussian & 1 & 0 & 16.38 & 15.76 & 14.52 & 14.08 & 13.22\\
$t$ & 1 & 0 & 16.15 & 15.51 & 14.00 & 13.43 & 12.32\\
$t$ & 5 & 0 & 13.61 & 12.66 & 10.96 & 10.40 & 9.34\\
skew $t$ & 1 & 0 & 9.24  & 7.27 & 4.13  & 3.27  & 1.96\\
skew $t$ & 5 & 0 & 15.81 & 14.46 & 11.57 & 10.57 & 8.60\\
$t$ & 1 & 0.9 & 5.52  & 3.58  & 1.77  & 1.47  & 1.10\\
$t$ & 5 & 0.9 & 5.98  & 4.27  & 2.41  & 2.03  & 1.49\\
 skew $t$ & 1 & 0.9 & {\bf 4.91}  & {\bf 3.16} & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
skew $t$ & 3 & 0.9 & 5.58 & 3.78 & 1.93& 1.58& 1.11\\
\hline
    \end{tabular}
  \end{table}
\end{frame}
